---
title: "06_update"
author: "randy"
date: '2022-04-19'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```


```{r include=FALSE}
library(tidyverse)
library(magrittr)

library(MASS)
library(here)

library(brokenstick)
library(gamlss)
library(lme4)

library(data.table)
library(gtsummary)
library(flextable)
library(splines)


library(ggridges)
```


1. Redo table 1 with final sample  

```{r include=FALSE}
table0 <- here::here("data", "epic_table1.csv") %>%
  read.csv(row.names = 1) %>%
  unique() %>%
  dplyr::select(-id) %>%
  mutate(
    ethnic = case_when(ethnic == 1 ~ "Hispanic",
                     ethnic == 2 ~ "Non-Hispanic"),
    race = case_when(race == 1 ~ "White",
                     race != 1 ~ "Other"),
    sex = case_when(sex == "F" ~ "Female",
                    sex == "M" ~ "Male"),
    age_diff = age_max - age_min) %>% 
  dplyr::select(
         Genotype = genotype,
         Gender = sex,
         Race = race,
         Ethnicity = ethnic,
         "Visit number" = visitn,
         "Age at registration, years" = age_min, 
         "Age at end of follow-up, years" = age_max, 
         "Follow-up, years" = age_diff,
         "Height at registration, cm" = h_min,
         # "Height mean" = h_mean, 
         # "Weight mean" = w_mean,
         "Weight at registration, kg" = w_min) %>% 
  ## select all the variables for table1
  tbl_summary() %>%
  ## just display all the variables in one column
  modify_header(label = "**Variable**") %>%
  # update the column header
  bold_labels() %>%
  italicize_labels() %>%
  as_flex_table() %>%
  flextable::bold(part = "header") %>% 
  ## auto adjust the column widths 
  flextable::autofit()

table1 <- here::here("data", "epic_table1.csv") %>%
  read.csv(row.names = 1) %>%
  unique() %>%
  dplyr::select(-id) %>%
  mutate(
    ethnic = case_when(ethnic == 1 ~ "Hispanic",
                     ethnic == 2 ~ "Non-Hispanic"),
    race = case_when(race == 1 ~ "White",
                     race != 1 ~ "Other"),
    sex = case_when(sex == "F" ~ "Female",
                    sex == "M" ~ "Male"),
    age_diff = age_max - age_min) %>% 
  dplyr::select(group,
         Genotype = genotype,
         Gender = sex,
         Race = race,
         Ethnicity = ethnic,
         "Visit number" = visitn,
         "Age at registration, years" = age_min, 
         "Age at end of follow-up, years" = age_max, 
         "Follow-up, years" = age_diff,
         "Height at registration, cm" = h_min,
         # "Height mean" = h_mean, 
         # "Weight mean" = w_mean,
         "Weight at registration, kg" = w_min) %>% 
  ## select all the variables for table1
  tbl_summary(by = group) %>%
  ## just display all the variables in one column
  modify_header(label = "**Variable**") %>%
  # update the column header
  bold_labels() %>%
  italicize_labels() %>%
  as_flex_table() %>%
  flextable::bold(part = "header") %>% 
  ## auto adjust the column widths 
  flextable::autofit()
```


```{r}
table0
table1

## save pptx -----------------------------------------------
## flextable can be saved directly to powerpoints
flextable::save_as_pptx(
  table0, 
  path = "figure/01_table0.pptx")
flextable::save_as_pptx(
  table1, 
  path = "figure/01_table1.pptx")
```


2. Refine spline model for training dataset. 

Try spline model with 3 inner knots set at percentiles (quartiles) too 
and compare to the splines you had tried before. 

```{r}
train <- here::here("data", 
                    "epic",
                    "epic_clean_randy.csv") %>%
  read.csv(row.names = 1) %>% 
  mutate(id = as.factor(id)) %>%
  filter(group == "training")

test <- here::here("data", 
                   "epic",
                   "epic_clean_randy.csv") %>%
  read.csv(row.names = 1) %>%
  filter(group == "testing")

basetest <- test %>% 
  group_by(id) %>% 
  arrange(time) %>% 
  slice(1L) 

View(train)
```


```{r}
# lmeControl
ctrl <- lmeControl(opt = 'optim')

## the original mod2_2_4 in file "04_model_fit.Rmd"
fit2_0 <- lme(ht ~ bs(time, knots = c(10, 15), 
                      degree = 3,
                      intercept = FALSE) * sex +  genotype - 1,
              control = ctrl,
              random = ~ 1| id,
              data = train)

fit2_1 <- lme(ht ~ bs(time, knots = c(10, 12), 
                      degree = 3,
                      intercept = FALSE) * sex +  genotype - 1,
              control = ctrl,
              random = ~ 1| id,
              data = train)

fit2_2 <- lme(ht ~ bs(time, df = 5, 
                      degree = 3,
                      intercept = FALSE) * sex +  genotype - 1,
              control = ctrl,
              random = ~ 1| id,
              data = train)

fit3_0 <- lme(ht ~ bs(time, df = 6, 
                      degree = 3,
                      intercept = FALSE) * sex +  genotype - 1,
              control = ctrl,
              random = ~ 1| id,
              data = train)
```


```{r}
get_tag <- function(fit, data1 = train) {
  tidy <- broom.mixed::tidy(fit)
  glance <- broom.mixed::glance(fit)
  augment <- broom.mixed::augment(fit) %>%
    dplyr::select(time, .fitted, sex) %>%
    unique()
  
  plot1 <- ggplot() +
        geom_smooth(data = data1, 
                    aes(x = time, 
                        y = ht, 
                        group = sex, 
                        color = sex)) +
        geom_smooth(data = augment, 
                    aes(x = time, y = .fitted, 
                        group = sex),
                    linetype = 2,
                    size = 0.5) +
        xlab("Age") +
        theme_bw() +
    theme(legend.position = "none")


  return(list(tidy = tidy, 
              glance = glance, 
              augment = augment,
              plot = plot1))
}
```

```{r}
tag2_0 <- get_tag(fit = fit2_0)
tag2_1 <- get_tag(fit = fit2_1)
tag2_2 <- get_tag(fit = fit2_2)
tag3_0 <- get_tag(fit = fit3_0)

glance_all <- map_dfr(list(tag2_0, tag2_1,
                    tag2_2, tag3_0),
                  "glance") %>%
  mutate(model = c("fit2_0", "fit2_1", 
                   "fit2_2", "fit3_0")) %>%
  dplyr::select(model, everything())

glance_all
```

3. plot the predicted/expected means. 

```{r echo=FALSE, fig.height=9, fig.width=10, message=FALSE, warning=FALSE}
# plot_all <- map(list(tag2_0, tag2_1,
#                     tag2_2, tag3_0),
#                   "plot")
# unlist(plot_all)
library("gridExtra")
grid.arrange(tag2_0$plot, tag2_1$plot,
             tag2_2$plot, tag3_0$plot,
             ncol = 2, nrow = 2)
```

4. For brokenstick model try time for a anchoring of 10 and 12  

```{r}
# View(brokenstick)
# View(brokenstick::brokenstick_bridge)

bks1 <- brokenstick(formula = ht ~ time | id,
                    data = train,
                    knots = c(10, 15),
                    method = "lmer")

bks10 <- brokenstick(formula = ht ~ time | id,
                    data = train,
                    subjid = train$id,
                    knots = c(10, 15))

est1_knots <- predict(bks1, 
                     x = "knots", 
                     group = train$id)
est1_all <- predict(bks1, 
                   group = train$id)


bks2 <- brokenstick(formula = ht ~ time | id,
                    data = train,
                    subjid = train$id,
                    knots = c(10, 12))
est2_knots <- predict(bks2, 
                     x = "knots", 
                     group = train$id)
est2_all <- predict(bks2, 
                   group = train$id)


bks31 <- brokenstick(formula = ht ~ time | id,
                    data = train,
                    subjid = train$id,
                    knots = c(10, 12, 15))

bks32 <- brokenstick(formula = ht ~ time | id,
                    data = train, 
                    knots = c(10, 12, 15),
                    seed = 555)

est31_knots <- predict(bks31, 
                     x = "knots", 
                     group = train$id)

est31_all <- predict(bks31, 
                   group = train$id)

est32_knots <- predict(bks32, 
                     x = "knots", 
                     group = train$id)
est32_all <- predict(bks32, 
                   group = train$id)
```

the biometrics paper,


```{r}
summary(bks31)
summary(bks32)
```

## the estimated c(10, 12) and c(10, 15)

```{r fig.height=5, fig.width=5}
plot_bks1 <- ggplot() +
  geom_smooth(data = train, 
              aes(x = time, y = ht, 
                  group = sex),
              alpha = 0.3,
              size = 1,
              color = "darkgreen") +
  geom_smooth(data = est1_all, 
              aes(x = time, y = .pred, 
                  group = sex),
              color = "brown",
              linetype = 2,
              size = 0.5) +
  geom_smooth(data = est2_all, 
                    aes(x = time, y = .pred, 
                        group = sex),
              color = "magenta",
              linetype = 4,
              size = 0.5) +
  xlab("Age") +
  theme_bw()

plot_bks1
```


The estimated for 2 **c(10, 12)** and 3 c(10, 12, 15) knots

## try c(10, 13), c(10, 14) ---------------------

```{r fig.height=5, fig.width=5}
plot_bks2 <- ggplot() +
  geom_smooth(data = train, 
              aes(x = time, y = ht, 
                  group = sex),
              alpha = 0.3,
              size = 1,
              color = "darkgreen") +
  geom_smooth(data = est32_all, 
              aes(x = time, y = .pred, 
                  group = sex),
              color = "brown",
              linetype = 2,
              size = 0.5) +
  geom_smooth(data = est2_all, 
                    aes(x = time, y = .pred, 
                        group = sex),
              color = "magenta",
              linetype = 4,
              size = 0.5) +
  xlab("Age") +
  theme_bw()

plot_bks2

```


```{r}
Kk <- c(5, 10, 12, 15)

Kplot <- function(K,
                  seed = 555,
                  form = "ht ~ time | id") {
  
  bks3n <- brokenstick(formula = as.formula(form),
                    data = train, 
                    knots = Kk,
                    method = "lmer")
  # ?brokenstick

  est_knots <- predict(bks3n, 
                     x = "knots", 
                     group = train$id)
  est_all <- predict(bks3n, 
                   group = train$id)
  
  plot_bks <- ggplot() +
    geom_smooth(data = train, 
                aes(x = time, y = ht, 
                    group = sex),
                alpha = 0.3,
                size = 1,
                color = "darkgreen") +
    geom_smooth(data = est_all, 
                aes(x = time, y = .pred, 
                    group = sex),
                color = "brown",
                linetype = 2,
                size = 0.5) +
    xlab("Age") +
    theme_bw()
  
  return(list(model = bks3n,
              est.knot = est_knots,
              est.all = est_all,
              plot = plot_bks))
}

bksn <- Kplot(K = Kk)
bksn$plot
```

```{r fig.height=3, fig.width=5}
tag2_1 <- get_tag(fit = fit2_1)

bks2_1 <- Kplot(K = c(10, 12))
bks2_1$plot

grid.arrange(tag2_1$plot,
             bks2_1$plot,
             ncol = 2)
```


```{r fig.height=3, fig.width=5}
plot <- ggplot() +
  geom_line(data = train, 
              aes(x = time, 
                  y = ht, 
                  group = id),
            color = "grey",
              alpha = 0.3,
              size = 0.2) +
  geom_smooth(data = bks2_1$est.all, 
                aes(x = time, y = .pred),
                color = "brown",
                linetype = 2,
                size = 0.5) +
        geom_smooth(data = tag2_1$augment, 
                    aes(x = time, y = .fitted, 
                        group = sex),
                    linetype = 3,
                    color = "purple",
                    size = 0.5) +
  xlab("Age") +
  theme_bw() 

plot + facet_wrap("sex")
```



5. Make progress, as able with the latex file of the paper. 

Start drafting a description of the dataset using Moss and Rice’s papers: 
one paragraph to describe the study and a second to include descriptive statistics.   
 

```{r fig.height=5, fig.width=5, include=FALSE}
# install.packages("IrregLong")
library(IrregLong)

abacus.plot(n = 10,
  time = "time",
  id = "id",
  data = train,
  tmin = 0,
  tmax = 16,
  # xlab.abacus = "Time",
  # ylab.abacus = "Subject",
  pch.abacus = 16,
  col.abacus = 1)
```


```{r}
set.seed(555)
f_train <- train %>%
  filter(sex == "F")

m_train <- train %>%
  filter(sex == "M")

sub_id <- c(sample(f_train$id, size = 10, replace = FALSE),
            sample(m_train$id, size = 10, replace = FALSE))

subtrain <- train %>%
  filter(id %in% sub_id) 

```


```{r fig.height=3, fig.width=5}
abacus0 <- ggplot(subtrain, 
       aes(y = as.factor(id), x = time, 
           group = sex,
           fill = sex, color = sex)) +
  geom_point(alpha = 0.5, size = 3) + 
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("Time") +
  ylab("ID")

abacus0

ggsave("figure/abacus.png")
```


```{r fig.height=5, fig.width=5}
ind_plot <- 
  ggplot(subtrain, 
         aes(y = ht, 
             x = time, 
             group = id,
             col = sex)) +
  geom_point(size = 0.5) +
  geom_line() +
  theme_bw() +
  theme(legend.position = "none") +
  facet_wrap(id ~ sex, nrow = 5)

ind_plot

ggsave("figure/sample_plot.png")
```



Dynamic Predictions in Bayesian Functional Joint Models for Longitudinal and Time-to-Event Data: An Application to Alzheimer’s Disease


