---
title: "01_initiator"
author: "randy"
date: "2022-07-21"
output: html_document
---

## prepare

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install.packages(c("devtools", 
#                    "roxygen2", 
#                    "testthat", 
#                    "usethis", 
#                    "knitr"))

library(tidyverse)
library(testthat)
library(usethis)
library(devtools)

load_all()
library(plmlmm)

# ?testthat::test_file
# ?devtools::test
# path <- testthat_example("success")
# test_file(path)
# test_file(path, reporter = "minimal")
```

## initiation

```{r}
packageVersion("devtools")
## to create a package
create_package("~/Desktop/project/plmlmm")
use_mit_license()

## look for R home directory
R.home()
## look for R code base
system.file()
## look at the files in the same directory
here::here() %>%
  list.files()
# list.dirs()
# here::here() %>%
#   list.dirs(full.names = FALSE,
#             recursive = TRUE)
```


## use_git

```{r}
## use git and github for the package
use_git()
usethis::use_git_remote("origin", 
                        url = NULL, 
                        overwrite = TRUE)
```



## use package

```{r}
use_package("brokenstick", type = "Depends")
use_package("broom.mixed", type = "Depends")

use_package("dplyr", type = "Depends")

use_package("forcats", type = "Depends")

use_package("gamlss", type = "Depends")

use_package("here", type = "Depends")

use_package("janitor", type = "Depends")
use_package("JMbayes", type = "Depends")

use_package("lme4", type = "Depends")

use_package("matrixcalc", type = "Depends")
use_package("MASS", type = "Depends")

use_package("nlme", type = "Depends")

use_package("tidyverse", type = "Depends")
use_package("tidyr", type = "Depends")
use_package("tibble", type = "Depends")

use_package("purrr")

use_package("readr")
use_package("rjags", type = "Depends")

use_package("stringr")
use_package("splines", type = "Depends")
use_package("shiny", type = "Depends")

devtools::install_github("Goodgolden/plmlmm")
```


## use_r & use_test

```{r}
# use_r("00_utils")
# use_r("01_plm")
# use_r("02_lmm")
# use_r("03_simulation")
# use_r("04_test")
# 
# use_test("01_plm")
# use_test("02_lmm")
# use_test("03_simulation")
```


## environment

```{r}
load_all()
exists("plm",
       # where = globalenv(), 
       inherits = FALSE)
#> [1] FALSE

# ?exists Environment Access
env <- environment(plm)
exists("plm", 
       where = env,
       # where = globalenv(), 
       inherits = FALSE)
```


## documents

Use `Ctrl + .` in RStudio and start typing 'DESCRIPTION' to activate a helper that makes it easy to open a file for editing. 

In addition to a filename, your hint can be a function name. 

```{r}
## document the description
document()
## always check 
## it's good to establish the habit of checking this often.
check()
```


```{r}
load_all()
# rm(list = c("distance_df", "linear_impute", "match", "norm2", "not_all_na", "not_any_na",
#   "people_like_me", "plm_ind_plot", "predict_gamlss", "single_df", "soft_fun"))
## `covr` package to track 
## what proportion of your package’s source code
## is exercised by the tests.
# install.packages("covr")
library("covr")
report()
test()
```

**Cmd + Shift + T (macOS)** for test

## readme

```{r}
## add description files to the package
# usethis::use_tidy_description()


## add datasets to the package
usethis::use_data(test, test_baseline,
                  all, margin_mean,
                  train, train_baseline,
                  demog, epic,
                  internal = TRUE,
                  overwrite = TRUE)

## build up a readme file and push into github
use_readme_rmd()
build_readme()

usethis::use_roxygen_md()
document()

```


Here is a typical sequence of calls when using devtools for package development:

- Edit one or more files below R/.
- document() (if you’ve made any changes that impact help files or  NAMESPACE)
- load_all()
- Run some examples interactively.
- test() (or test_file())
- check()


```{r}
load_all()
plmlmm:::run_shiny()
```

